\section{Implementation}
This section summarizes the all sections above in terms of work needed.  This
aims to be a giant to-do list.

There are essentially two independent modules:
\begin{description}
\item{\bf Convert arbitrary elasticity tensors into pattern parameters}\\
{\it Input:} An array of elasticity tensors.

{\it Output:} An array of pattern parameters (and fitting errors).  A bound for
all possible material properties that can be generated with a given type of patterns.

{\it Validation:} If homogenization has closed formula, we should validate it
with simulation.

{\it Physical experiments:}  Validate homogenization formula on simple patterns.\\

Approach 1: Look-up Table
\begin{itemize}
\item For microstructures with closed formulas for homogenized elasticity
    tensors, sample the space of all possible pattern parameter combinations.
\item For microstructures without closed formulas, use the representative volume
method (numerical coarsening) to find them.  I.e., for all possible pattern
parameters, generate a cube of microstructures and simulate with periodic
boundary conditions its behavior using a very fine mesh or mesh-free FEM.
Use \cite{Kharevych2009} to obtain the homogenized elasticity tensors.
\item Build a table that supports fast lookup of pattern parameters with the
    elasticity tensor as key. Return an interpolation of closest entries using a
    reasonable distance measure.
\item Optionally run a few iterations of the optimization described in Approach
      2 to improve the fit.
\end{itemize}

Approach 2: Direct Optimization of Material Parameters.
\begin{itemize}
\item For microstructures with closed formulas for homogenized elasticity
    tensors, directly optimize the Frobenius distance from the target elasticity
    tensor with a nonlinear optimization algorithm, e.g. Newton's method.
\end{itemize}

\item{\bf Generate microstructure geometry}\\
{\it Input:} An array of pattern parameters.

{\it Output:} Microstructure mesh

{\it Validation:} Simulate the resulting microstructure to see if its behaviors
are expected.

{\it Physical experiments:} Validate generated patterns in experiments.

Approach 1: tile pattern for each element and stitch.
\begin{itemize}
\item Write code to generate regular tiled patch for each element.
\item Write code to stitch patches together without significantly effect the material
properties.
\item Verify stitching does not introduce artifacts globally.
\end{itemize}

Approach 2: tracing dual graph and pack holes.
\begin{itemize}
\item Implement or borrow code for streamline tracing.  Make sure streamlines
form a network.
\item Write code to procedurally generate holes at each stream line
intersection.
\item Verify the resulting mesh is a valid microstructure.
\end{itemize}
\end{description}



There is also work involved in generating results and validations.
\begin{itemize}
\item Mesh-free 3D simulation.
\item Material property painting program.  It would be handy to generate
inputs.
\item Code for computing homogenized material property from
\cite{Kharevych2009}.  This would be helpful to generate elasticity tensor from
load specification and deformation field.
\end{itemize}



